### 2013-01-14 report/content/listings.tex
Finished writing the paper

\lstinputlisting[label=cp:typ,language=C++,caption={Templates and static typing in C++}]{../C++/Types.cpp}
\lstinputlisting[label=oc:typ,caption={Objective-C using Dynamic and Static typing}]{../objC/Types.m}

\lstinputlisting[label=cp:acc,language=C++,caption={Access rights in C++}]{../C++/Access.cpp}
\lstinputlisting[label=oc:acc,caption={Access rights in Objective-C}]{../objC/Access.m}

\lstinputlisting[label=cp:pol,language=C++,caption={Polymorphism in C++}]{../C++/Polymorphism.cpp}
\lstinputlisting[label=oc:pol,caption={Polymorphism in Objective-C}]{../objC/Polymorphism.m}

\lstinputlisting[label=cp:inh,language=C++,caption={Inheritance in C++}]{../C++/Inheritance.cpp}
\lstinputlisting[label=cp:min,language=C++,caption={Multiple inheritance in C++}]{../C++/MultipleInheritance.cpp}
\lstinputlisting[label=oc:inh,caption={Inheritance in Objective-C}]{../objC/Inheritance.m}

\lstinputlisting[label=cp:con,language=C++,caption={(De)Constructors in C++}]{../C++/Constructors.cpp}
\lstinputlisting[label=oc:con,caption={(De)Constructors in Objective-C}]{../objC/Constructors.m}

\lstinputlisting[label=oc:pro,caption={A protocol in Objective-C}]{../objC/Protocol.m}
\lstinputlisting[label=oc:cat,caption={A category in Objective-C}]{../objC/Categories.m}

\lstinputlisting[label=cp:for,language=C++,caption={Message forwarding in C++}]{../C++/Forwarding.cpp}
\lstinputlisting[label=oc:for,caption={Message forwarding in Objective-C}]{../objC/Forwarding.m}

\lstinputlisting[label=oc:ref,caption={Changing and adding methods at runtime in Objective-C}]{../objC/Reflection.m}

\lstinputlisting[label=cp:met,language=C++,caption={Method overloading in C++}]{../C++/MethodOverloading.cpp}

\lstinputlisting[label=cp:nam,language=C++,caption={Namespaces in C++}]{../C++/Namespaces.cpp}

\lstinputlisting[label=cp:abs,language=C++,caption={Abstract class in C++}]{../C++/Abstract.cpp}
\lstinputlisting[label=oc:abs,caption={Replicating abstract classes in Objective-C}]{../objC/Abstract.m}

\lstinputlisting[label=cp:fri,language=C++,caption={Friend classes in C++}]{../C++/Friend.cpp}
\lstinputlisting[label=cp:inn,language=C++,caption={Inner classes in C++}]{../C++/InnerClass.cpp}